---
title: "PED Mis. Sample Tables - Report"
header-includes: 
   \usepackage{graphicx}
   \usepackage{fancyhdr}
   \pagestyle{fancy}
   \setlength\headheight{28pt}
   \fancyhead[L]{\includegraphics[width=0.5cm]{mobovidalogo.png}}
   \fancyfoot[LE,RO]{CL}
author: "Chloe Li"
date: "September 6, 2016"
output:
  pdf_document: 
    highlight: monochrome
  html_document:
    highlight: pygments
    theme: cerulean
classoption: landscape
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
setwd("/Users/chloeli/Documents/01. PriceElasticity_CL/004.Reports")

TopSellers_PED <- read.csv("PED_TopSellers.csv")
PED_byBrand <- read.csv("PED_byBrand.csv")
PED_allproduct <- read.csv("PED_AllProduct.csv")

pacman::p_load('ReporteRs',"knitr","xtable")

TopSellers_PED$X <- NULL
PED_byBrand$X <- NULL
PED_allproduct$X <- NULL
PED_byBrand$Change_price <- NULL
PED_byBrand$Change_Avgqty <- NULL
      
```

## Objectives
- *How should we project the change in demand when we intend to change a certain product's price?*

## Introduction
- This report intends to show 3 sample PED tables
- 1) PED calculation for all product
- 2) PED calculation for top selling product with high average demand per day
- 3) PED calcualtion for all product by brand/model

\pagebreak

## Table 1 - PED calculation for all product

```{r, echo=FALSE, message=FALSE, warning=FALSE}

kable(PED_allproduct[1:30,])


```

\pagebreak

## Table 2 - PED calculation for top selling product with high average demand per day

```{r, echo=FALSE, message=FALSE, warning=FALSE}

kable(TopSellers_PED)

```

\pagebreak

## Table 3 - PED calcualtion for all product by brand/model

```{r, echo=FALSE, message=FALSE, warning=FALSE}

kable(PED_byBrand[1:34,])

```